---
title: "Cleaning template - COVID-19 Behind Bars Historical Data Cleaning"
author: "Hope Johnson"
date: "10/26/2020"
output: html_document
---

```{r package setup, include=FALSE}
##Define package list
Packages<-c("tidyverse")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
```

## Purpose

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Chase Hommeyer, Grace DiLaura, and Kalind Parish. 


## Load inputs

Input files: 

* Utilities script
* Historical data to clean
* Date range to clean 
* Facility name look-up table 

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code", "historical", "historical-data")
data_folder <- file.path("data", "xlsx")

##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
##Facility name look-up table
f <- read_xlsx(file.path(base_path, data_folder,"facility_spellings_091820.xlsx"))
##Historical data file
historical_data <- file.path(base_path, data_folder, "Covid Custody Project_100720.xlsx")
```


