---
title: "data cleaning Oklahoma"
author: "Kevin Zhao"
date: "4/15/2021"
output: pdf_document
---

```{r package setup, include=FALSE}
rm(list=ls())
##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr", "skimr", "plotly")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
options(skimr_strip_metadata = FALSE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
library(behindbarstools)
help(package=behindbarstools)
##Define state
state_toclean <- "Oklahoma"
state_abbrev <- behindbarstools::translate_state(state_toclean, reverse = TRUE)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Erika Tyagi, Chase Hommeyer, Grace DiLaura,  and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 

## Load inputs

Input files: 

* Utilities script
* Historical data

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code",  "historical-data")
data_path <- file.path(base_path, "data", "inputs")
##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```

Load in all the historical data for this state!

```{r load data, message=FALSE}
df <- load_data(data_path, 
                "11420",
                filter_state = state_toclean) 
df_typed <- type_convert(df)
```

Fill in missing values of `Staff.Deaths` with non-missing values of `Staff.Death`, when those exist. Do the same with `Resident.Death`/`Resident.Deaths`. 

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) %>% # rm 100% missing cols 
  mutate(Zipcode = as.character(Zipcode),
         County.FIPS = as.character(County.FIPS))
# if there are any other similarly-named columns in the results below, pay attention!
df_out %>%
  select(contains("Staff")) %>% 
  names()
df_out %>%
  select(contains("Resident")) %>% 
  names()
```

```{r create date var}
df_out <- df_out %>%
  mutate(Date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r facility name cleaning checks}
#renaming observations in original df_out
df_mid <- df_out %>% 
  mutate(Name = str_replace(Name, "Correctiona ", "Correctional "))

df_cln <- behindbarstools::clean_facility_name(df_mid, debug = TRUE) 

#checking again
df_cln %>%
  filter(name_match == FALSE) %>% 
  select(scrape_name_clean) %>%
  unique()
```

Remove rows when we didn't get a name at all.
```{r}
## Getting rid of facilities named "-" 
## mutating because "North Fork Correctional Center- Unit is definitely J unit 
df_cln_out <- df_cln %>% 
  filter(Name != "-") %>% 
  filter(!is.na(Name)) %>%
  mutate(Name = ifelse(Name == "NORTH FORK CORRECTIONAL CENTER- UNIT", 
                       "NORTH FORK CORRECTIONAL CENTER J UNIT", 
                       Name)) %>%
  select(-c(jurisdiction_scraper, scrape_name_clean, federal_bool, Facility.ID, xwalk_name_clean,
            Jurisdiction, name_match))

df_cln_cln_out <- behindbarstools::clean_facility_name(df_cln_out, debug = TRUE) 

#checking again
df_cln_cln_out %>%
  filter(name_match == FALSE) %>% 
  select(scrape_name_clean) %>%
  unique()
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_cln_cln_out, Date, Name))
see_if(nrow(df_cln_cln_out) == nrow(distinct(df_cln_cln_out, Name, Date)))
# if there are two values, sum them 
df_comb <- df_cln_cln_out %>% 
  behindbarstools::group_by_coalesce(., Name, Facility.ID, Date, 
                                     .ignore = "scrape_name_clean",
                                     .method = "sum") 
assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, Date)))
```

Assertions / sanity checks ! 
```{r}
# make sure there's only 1 clean name per facility ID! 
df_comb %>% 
  group_by(Facility.ID) %>% 
  summarise(n_name = n_distinct(Name)) %>% 
  filter(n_name > 1)
# make sure there's only 1 facility ID per clean name!
df_comb %>% 
    group_by(Name) %>% 
    summarise(n_fac_ID = n_distinct(Facility.ID)) %>% 
    filter(n_fac_ID > 1)
```

Filter down and re-order columns in order to row bind them to latest data.

```{r}
df_hist <- behindbarstools::reorder_cols(df_comb)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Residents.Deaths = Resident.Deaths) %>%
  select(-c(Website, Resident.Deaths, Count.ID, Facility)) 
df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE) %>%
  select_if(~sum(!is.na(.)) > 0) # rm all NA
```

Make plots for the data, and take a look at any weirdness. 

```{r plot cases/deaths}
res_conf_stacked <- df_hist_final %>% 
  mutate(Facility.ID = as.character(Facility.ID)) %>%
  arrange(Facility.ID) %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = Residents.Confirmed, 
             group = Facility.ID)) +
  theme(legend.position = "none") + 
  geom_line(aes(colour = Name), position = "stack", alpha = 0.4, size = 0.5) + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y") 
ggplotly(res_conf_stacked)
res_death_stacked <- df_hist_final %>% 
  mutate(Facility.ID = as.character(Facility.ID)) %>%
  arrange(Facility.ID) %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = Residents.Deaths, 
             group = Facility.ID)) +
  theme(legend.position = "none") + 
  geom_line(aes(colour = Name), position = "stack", alpha = 0.4, size = 0.5) + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y") 
ggplotly(res_death_stacked)
# create a testing file 
testing_cases <- flag_noncumulative_cases(df_hist_final, Name)
testing_deaths <- flag_noncumulative_deaths(df_hist_final, Name, Residents.Deaths)
# lag cases overall 
all_cases <- testing_cases %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_cases, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
         scale_x_date(date_minor_breaks = "1 month", date_labels = "%m/%y", 
                      date_breaks = "1 month") + 
    labs(x = "Date",
      y = "lag_change_cases")
ggplotly(all_cases)
# lag deaths overall
all_deaths <- testing_deaths %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_deaths, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
    scale_x_date(date_labels = "%m") + 
    labs(x = "Date",
      y = "lag_change_deaths")
ggplotly(all_deaths)
```

Make facility-specific plots, and take a look at any weirdness.

```{r}
# will throw a warning if path already exists
plot_path <- file.path(base_path, "plots", state_abbrev)
dir.create(file.path(plot_path))
dir.create(file.path(plot_path, "cases"))
dir.create(file.path(plot_path, "deaths"))
# plot lag counts by facility
lag_case_plots <- plot_lags(testing_cases, "Date", 
                            y_var = "lag_change_cases",
                            grp_var = Name)
# lag_case_plots$plot # view them by un-commenting this
## save lag case plots
for (i in 1:nrow(lag_case_plots)){
  facility_name <- lag_case_plots$Name[[i]]
  ggsave(paste0(facility_name, "_LagChangeCases.png"), lag_case_plots$plot[[i]],
         path = file.path(base_path, "plots", state_abbrev, "cases"))
}
lag_death_plots <- plot_lags(testing_deaths, "Date", 
                            y_var = "lag_change_deaths",
                            grp_var = Name)
# save lag death plots
for (i in 1:nrow(lag_death_plots)){
  facility_name <- lag_death_plots$Name[[i]]
  ggsave(paste0(facility_name, "_LagChangeDeaths.png"), lag_death_plots$plot[[i]],
         path = file.path(base_path, "plots", state_abbrev, "deaths"))
}
```


Find date spans / week spans with no data. In instances where the count went down by one, it could be that a PDF was misread. 

```{r}
dates <- df_hist_final %>%
  arrange(Date) %>%
  count(Date)
dates
ggplot(data = dates, 
       aes(x = Date, y = n)) +
  geom_bar(stat="identity") +
  labs(x = "Date",
    y = "n instances")
```

Marshall Project / AP data comparison.

YIKES ! Data is crazy crazy. Not integrating this one. 

```{r}
mpap <- read_mpap_data(all_dates = TRUE) %>%
  filter(State == state_toclean)
ourdat_tocompare <- df_hist_final %>% 
  filter(Jurisdiction == "state") %>%
  group_by(Date) %>%
  summarise(Residents.Confirmed = sum_na_rm(Residents.Confirmed),
            Residents.Deaths = sum_na_rm(Residents.Deaths),
            Residents.Recovered = sum_na_rm(Residents.Recovered),
            Staff.Recovered = sum_na_rm(Staff.Recovered))
  
comparison <- ourdat_tocompare %>%
  full_join(mpap, by = "Date", suffix = c("_cbb", "_mpap")) %>%
  arrange(Date) %>%
  select(-State) %>% 
  select(Date, 
         starts_with("Residents.Confirmed"), 
         starts_with("Residents.Deaths"),
         starts_with("Residents.Recovered"),
         starts_with("Staff.Recovered"))
```

Fix issues from facility anomaly-checking:

```{r}
df_hist_final_edit <- df_hist_final %>%
  
```

Prep pre-November data for writing to the server. 

```{r}
df_hist_towrite <- prep_server_data(df_hist_final, state_abbrev)
skim(df_hist_towrite) # double-check everything in the data 
```

Write pre-November data to the server. 

```{r}
srvr_outfile_name <- glue('1-pre-november-{state_abbrev}.csv')
write_csv(df_hist_towrite, file.path(base_path, "data", "pre-nov", srvr_outfile_name))
```