---
title: "Clean Tennessee - COVID-19 Behind Bars Historical Data Cleaning"
author: "Hope Johnson & Dara Tan"
date: "4/01/21"
output: html_document
---

```{r package setup, include=FALSE}
rm(list=ls())
##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr", "skimr", "plotly")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
options(skimr_strip_metadata = FALSE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
library(behindbarstools)
help(package=behindbarstools)

##Define state
state_toclean <- "Tennessee"
state_abbrev <- behindbarstools::translate_state(state_toclean, reverse = TRUE)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Erika Tyagi, Chase Hommeyer, Grace DiLaura,  and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 

## Load inputs

Input files: 

* Utilities script
* Historical data

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code", "historical-data")
data_path <- file.path(base_path, "data", "inputs")
##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```

Load in all the historical data for this state!

```{r load data, message=FALSE}
df <- load_data(data_path, 
                "11420",
                filter_state = state_toclean) 
df_typed <- type_convert(df)
```

Fill in missing values of `Staff.Deaths` with non-missing values of `Staff.Death`, when those exist. Do the same with `Resident.Death`/`Resident.Deaths`. 

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) %>% # rm 100% missing cols 
  mutate(Zipcode = as.character(Zipcode),
         County.FIPS = as.character(County.FIPS))

# if there are any other similarly-named columns in the results below, pay attention!
df_out %>%
  select(contains("Staff")) %>% 
  names()

df_out %>%
  select(contains("Resident")) %>% 
  names()
```

```{r}
df_out %<>% 
  mutate(# Staff.Deaths =
           # behindbarstools::coalesce_with_warnings(Staff.Deaths, Staff.Death),
         Resident.Deaths =
           behindbarstools::coalesce_with_warnings(Resident.Deaths,
                                                   Resident.Death))
```

```{r create date var}
df_out <- df_out %>%
  mutate(Date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r standardize facility names}
df_mid <- behindbarstools::clean_facility_name(df_out, debug = TRUE) 
```

```{r filter out federal facilities}
df_mid <- df_mid %>%
  filter(federal_bool == FALSE)
```

Show instances where the  didn't identify a clean name. 

- Stop and add these to facility spellings and/or facility data, here: https://docs.google.com/spreadsheets/d/1tAhD-brnuqw0s55QXM-xYLPsyD-rNrqHbAVIbxSOMwI/edit#gid=0

- Reference the current working versions of both sheets, available here: https://github.com/uclalawcovid19behindbars/facility_data/tree/master/data 

```{r facility name cleaning checks}
# check if it's in fac_data
df_mid %>%
  filter(name_match == FALSE) %>% 
  select(scrape_name_clean) %>%
  unique()
```

Remove rows when we didn't get a name at all.
```{r}
df_mid <- df_mid %>%
  filter(!is.na(Name))
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_mid, Date, Name))
see_if(nrow(df_mid) == nrow(distinct(df_mid, Name, Date)))

# if there are two values, sum them 
df_comb <- df_mid %>% 
  behindbarstools::group_by_coalesce(., Name, Facility.ID, Date, 
                                     .ignore = "scrape_name_clean",
                                     .method = "sum") 

assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, Date)))
```

Assertions / sanity checks ! 
```{r}
# make sure there's only 1 clean name per facility ID! 
df_comb %>% 
  group_by(Facility.ID) %>% 
  summarise(n_name = n_distinct(Name)) %>% 
  filter(n_name > 1)

# make sure there's only 1 facility ID per clean name!
df_comb %>% 
    group_by(Name) %>% 
    summarise(n_fac_ID = n_distinct(Facility.ID)) %>% 
    filter(n_fac_ID > 1)
```
Remove "statewide" counts on 

```{r}
df_rmd <- df_comb %>%
  filter(Name != "STATEWIDE")
```

Filter down and re-order columns in order to row bind them to latest data.

```{r}
df_hist <- behindbarstools::reorder_cols(df_rmd)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Residents.Deaths = Resident.Deaths) %>%
  select(-c(Website, Resident.Deaths, Resident.Death, Count.ID, Facility)) 

df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE) %>%
  select_if(~sum(!is.na(.)) > 0) # rm all NA
```

Make plots for the data, and take a look at any weirdness. 

```{r plot cases/deaths}
res_conf_stacked <- df_hist_final %>% 
  mutate(Facility.ID = as.character(Facility.ID)) %>%
  arrange(Facility.ID) %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = Residents.Confirmed, 
             group = Facility.ID)) +
  theme(legend.position = "none") + 
  geom_line(aes(colour = Name), position = "stack", alpha = 0.4, size = 0.5) + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y") 
ggplotly(res_conf_stacked)

res_death_stacked <- df_hist_final %>% 
  mutate(Facility.ID = as.character(Facility.ID)) %>%
  arrange(Facility.ID) %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = Residents.Deaths, 
             group = Facility.ID)) +
  theme(legend.position = "none") + 
  geom_line(aes(colour = Name), position = "stack", alpha = 0.4, size = 0.5) + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y") 
ggplotly(res_death_stacked)


# create a testing file 
testing_cases <- flag_noncumulative_cases(df_hist_final, Name)
testing_deaths <- flag_noncumulative_deaths(df_hist_final, Name, Residents.Deaths)

# lag cases overall 
all_cases <- testing_cases %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_cases, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
         scale_x_date(date_minor_breaks = "1 month", date_labels = "%m/%y", 
                      date_breaks = "1 month") + 
    labs(x = "Date",
      y = "lag_change_cases")

ggplotly(all_cases)

# lag deaths overall
all_deaths <- testing_deaths %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_deaths, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
    scale_x_date(date_labels = "%m") + 
    labs(x = "Date",
      y = "lag_change_deaths")
ggplotly(all_deaths)
```

```{r COMMENTS BY DARA}
ggplotly(all_cases)
  # 2020-05-07 -> change = -38 for TURNEY CENTER INDUSTRIAL COMPLEX
    # 38 on 2020-05-06, 0 on 2020-05-07 and 38 on 2020-05-08
  # checked using Way Back Machine -> numbers match -> issue can be ignored
```

Make facility-specific plots, and take a look at any weirdness.

```{r}
# will throw a warning if path already exists
plot_path <- file.path(base_path, "plots", state_abbrev)
dir.create(file.path(plot_path))
dir.create(file.path(plot_path, "cases"))
dir.create(file.path(plot_path, "deaths"))
dir.create(file.path(plot_path, "staff confirmed"))

# CREATE PATH MANUALLY

# plot lag counts by facility
lag_case_plots <- plot_lags(testing_cases, "Date", 
                            y_var = "lag_change_cases",
                            grp_var = Name)
# lag_case_plots$plot # view them by un-commenting this

## save lag case plots
for (i in 1:nrow(lag_case_plots)){
  facility_name <- lag_case_plots$Name[[i]]
  ggsave(paste0(facility_name, "_LagChangeCases.png"), lag_case_plots$plot[[i]],
         path = file.path(base_path, "plots", state_abbrev, "cases"))
}

lag_death_plots <- plot_lags(testing_deaths, "Date", 
                            y_var = "lag_change_deaths",
                            grp_var = Name)
# save lag death plots
for (i in 1:nrow(lag_death_plots)){
  facility_name <- lag_death_plots$Name[[i]]
  ggsave(paste0(facility_name, "_LagChangeDeaths.png"), lag_death_plots$plot[[i]],
         path = file.path(base_path, "plots", state_abbrev, "deaths"))
}
```

Investigating issues with the number of staff cumulative cases:

Remove any huge dips/jumps >10 cases: 

```{r}
df_hist_final_corrected <- df_hist_final %>%
  mutate(Residents.Confirmed  = ifelse((Name == "TURNEY CENTER INDUSTRIAL COMPLEX") & (Date == "2020-05-07"),
                                       NA,
                                       Residents.Confirmed),
         Residents.Confirmed  = ifelse((Name == "MARK LUTTRELL TRANSITION CENTER") & (Date == "2020-05-18"),
                                       NA,
                                       Residents.Confirmed),
        Staff.Confirmed = ifelse((Name == "SOUTH CENTRAL CORRECTIONAL FACILITY") & (Date == "2020-09-23"),
                             NA,
                             Staff.Confirmed),
       Staff.Confirmed = ifelse((Date == "2020-10-01"),
                             NA,
                             Staff.Confirmed),
        Staff.Confirmed = ifelse((Name %in% c("MORGAN COUNTY CORRECTIONAL COMPLEX",
                                              "LOIS M DERRY SPECIAL NEEDS FACILITY",
                                              "TURNEY CENTER INDUSTRIAL COMPLEX",
                                              "DEBRA K JOHNSON REHABILITATION CENTER"
                                              )) & 
                                   (Date == "2020-10-03"),
                             NA,
                             Staff.Confirmed),
         )
```

```{r}

staff_compare <- df_hist_final_corrected %>% 
  filter(Date >= as.Date("2020-09-29") & Date <= as.Date("2020-10-03")) %>% 
  group_by(Name) %>% 
  mutate(prev_conf = dplyr::lag(Staff.Confirmed, order_by = Date),
         change_conf = Staff.Confirmed - prev_conf) %>%
  ungroup() %>%
  select(Date, Name, Staff.Confirmed, change_conf) 
View(staff_compare)

staff_compare %>%
  group_by(Date) %>%
  summarise(count = sum_na_rm(Staff.Confirmed),
            n = n())
## seems like for 10/01, lots of facilities had counts that were not recorded before or after
```


```{r COMMENTS BY DARA}
# case plots ----

  # BLEDSOE COUNTY CORRECTIONAL COMPLEX
    # negative change: mid-05/20
testing_cases %>%
  filter(Name == "BLEDSOE COUNTY CORRECTIONAL COMPLEX",
         Date > "2020-05-10", Date < "2020-05-20") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-05-17; change = -1
    # checked using WBM -> numbers match positive + recovered + deaths

  # HARDEMAN COUNTY CORRECTIONAL FACILITY
    # negative change: late-05/20
testing_cases %>%
  filter(Name == "HARDEMAN COUNTY CORRECTIONAL FACILITY",
         Date > "2020-05-15", Date < "2020-05-25") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-05-21; change = -1
    # checked using WBM -> numbers match positive

  # LOIS M DERRY SPECIAL NEEDS FACILITY
    # negative change: late-09/20
testing_cases %>%
  filter(Name == "LOIS M DERRY SPECIAL NEEDS FACILITY",
         Date > "2020-09-20", Date < "2020-10-05") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-09-29; change = -1
      # 2020-09-29: WBM positive = 5 and recovered = 121 and deaths = 7, CBB = 134 -> FLAG

  # MARK LUTTRELL TRANSITION CENTER
    # negative change: mid-05/20
  ## REMOVED 2020-05-18 CASE COUNT (TO NA)
testing_cases %>%
  filter(Name == "MARK LUTTRELL TRANSITION CENTER",
         Date > "2020-05-10", Date < "2020-05-20") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-05-19; change = -10
      # 2020-05-18: WBM positive = 53 and recovered = 1, CBB = 55 -> FLAG
      # 2020-05-19: WBM positive = 53, CBB = 45 -> FLAG

  # MORGAN COUNTY CORRECTIONAL COMPLEX
    # negative change: mid-09/20
testing_cases %>%
  filter(Name == "MORGAN COUNTY CORRECTIONAL COMPLEX",
         Date > "2020-09-10", Date < "2020-09-20") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-09-16; change = -1
    # checked using WBM -> numbers match recovered

  # NORTHWEST CORRECTIONAL COMPLEX
    # negative change: late-05/20
testing_cases %>%
  filter(Name == "NORTHWEST CORRECTIONAL COMPLEX",
         Date > "2020-05-15", Date < "2020-05-25") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-05-21; change = -3
      # 2020-05-21: checked using WBM -> numbers match positive + recovered
      # 2020-05-22: WBM positive = 331 and recovered = 51, CBB = 379 -> FLAG

  # RIVERBEND MAXIMUM SECURITY INSTITUTION
    # negative change: mid-09/20
testing_cases %>%
  filter(Name == "RIVERBEND MAXIMUM SECURITY INSTITUTION",
         Date > "2020-09-10", Date < "2020-09-20") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-09-14; change = -1
    # checked using WBM -> numbers match positive + recovered

  # SOUTH CENTRAL CORRECTIONAL FACILITY
    # negative change: late-10/20
testing_cases %>%
  filter(Name == "SOUTH CENTRAL CORRECTIONAL FACILITY",
         Date > "2020-10-11", Date < "2020-10-20") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-10-18; change = -1
    # checked using WBM -> numbers match positive + recovered + deaths

  # *STATEWIDE*
    # negative change: late-09/20
testing_cases %>%
  filter(Name == "STATEWIDE",
         Date > "2020-09-20", Date < "2020-09-30") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-09-23; change = -1 -> flag issue

  # TROUSDALE TURNER CORRECTIONAL CENTER
    # negative change: late-07/20
testing_cases %>%
  filter(Name == "TROUSDALE TURNER CORRECTIONAL CENTER",
         Date > "2020-07-15", Date < "2020-07-25") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-07-21; change = -2
      # 2020-07-21: WBM recovered = 1378 and deaths = 3, CBB = 1383 -> FLAG

  # TURNEY CENTER INDUSTRIAL COMPLEX
    # negative changes: (1) early 04-20, (2) early-05/20, (3) early-05/20, (4) mid-05/20
## CHANGED 2020-05-07 Residents.Confirmed to NA
testing_cases %>%
  filter(Name == "TURNEY CENTER INDUSTRIAL COMPLEX",
         Date > "2020-04-05", Date < "2020-05-25") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: (1) 2020-04-09 = -1; (2) 2020-05-07 = -38;
    # (3) 2020-05-12 = -3; (4) 2020-05-21 = -1
      # (1) no WBM captures; (2) see chunk 15; (3) matches WBM positive + recovered;
      # (4) matches WBM positive + recovered for 21 but no captures for 22

  # WEST TENNESSEE STATE PENITENTIARY
    # negative change: late-05/20
testing_cases %>%
  filter(Name == "WEST TENNESSEE STATE PENITENTIARY",
         Date > "2020-05-15", Date < "2020-05-25") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-05-21; change = -1
      # 2020-05-21: matches WBM positive + recovered
      # 2020-05-22: no WBM captures

  # WHITEVILLE CORRECTIONAL FACILTY -> FLAG TYPO?
    # negative change: early-08/20

## NB: correcting the name of this in the crosswalk (facility is misspelled)! 

testing_cases %>%
  filter(Name == "WHITEVILLE CORRECTIONAL FACILTY",
         Date > "2020-08-05", Date < "2020-08-15") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-08-11; change = -1
      # 2020-08-11: WBM recovered = 41 and deaths = 2, CBB = 44 -> FLAG

  # WOMENS THERAPEUTIC RESIDENTIAL CENTER AND WOMENS THERAPEUTIC RESIDENTIAL CENTER
    # negative change: early-05/20
testing_cases %>%
  filter(Name ==
           "WOMENS THERAPEUTIC RESIDENTIAL CENTER AND WOMENS THERAPEUTIC RESIDENTIAL CENTER",
         Date > "2020-05-05", Date < "2020-05-15") %>%
  select(Date, previous_date_value_cases, lag_change_cases)
    # exact date: 2020-05-12; change = -1
    # checked using WBM -> numbers match positive

# no problematic death plots
```

Find date spans / week spans with no data. In instances where the count went down by one, it could be that a PDF was misread. 

```{r}
dates <- df_hist_final_corrected %>%
  arrange(Date) %>%
  count(Date)
dates

ggplot(data = dates, 
       aes(x = Date, y = n)) +
  geom_bar(stat="identity") +
  labs(x = "Date",
    y = "n instances")
```

```{r COMMENTS BY DARA}
df_hist_final_corrected %>%
  filter(Date > "2020-08-31") %>%
  count(Name)
# 2 peaks after August are due to 'STATEWIDE'
```

Marshall Project / AP data comparison.

```{r}
mpap <- read_mpap_data(all_dates = TRUE, window = 14) %>%
  filter(State == state_toclean)

ourdat_tocompare <- df_hist_final_corrected %>% 
  filter(Jurisdiction == "state") %>%
  group_by(Date) %>%
  summarise(Residents.Confirmed = sum_na_rm(Residents.Confirmed),
            Residents.Deaths = sum_na_rm(Residents.Deaths),
            Residents.Recovered = sum_na_rm(Residents.Recovered),
            Staff.Confirmed = sum_na_rm(Staff.Confirmed),
            Staff.Recovered = sum_na_rm(Staff.Recovered))

comparison <- ourdat_tocompare %>%
  full_join(mpap, by = "Date", suffix = c("_cbb", "_mpap")) %>%
  arrange(Date) %>%
  select(-State) %>% 
  select(Date, 
         starts_with("Staff.Confirmed"),
         starts_with("Residents.Confirmed"), 
         starts_with("Residents.Deaths"),
         starts_with("Residents.Recovered"),
         starts_with("Staff.Recovered"))
```

**POSSIBLY PROBLEMATIC (DIFFERENCE > 20)**

- 2020-10-07 Residents Confirmed: CBB = 5577; MPAP = 5555

- 2020-09-29 to 2020-09-31 Residents Recovered:
  + CBB on 29 and 31 = 5186
  + MPAP on 30 = 5254

- 2020-06-30 Staff Recovered: CBB = 122; MPAP = 93

- For every date starting 2020-06-30 that MPAP has data for Staff Recovered, CBB and MPAP differ by at least 20 and the difference gets larger over time.


Prep pre-November data for writing to the server. 

```{r}
df_hist_towrite <- prep_server_data(df_hist_final_corrected, state_abbrev)

skim(df_hist_towrite) # double-check everything in the data 
```

Write pre-November data to the server. 

```{r}
srvr_outfile_name <- glue('1-pre-november-{state_abbrev}.csv')

write_csv(df_hist_towrite, file.path(base_path, "data", "pre-nov", srvr_outfile_name))
# write_csv(df_hist_towrite, srvr_outfile_name)

sync_remote_files(srvr_outfile_name) # ignore this line if you're not hope!
```
