---
title: "Clean Arkansas - COVID-19 Behind Bars Historical Data Cleaning"
author: "Hope Johnson"
date: "2/9/21"
output: html_document
---

```{r package setup, include=FALSE}
rm(list=ls())
##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr", "skimr", "plotly")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
help(package=behindbarstools)

##Define state
state_toclean <- "Arkansas"
state_abbrev <- behindbarstools::translate_state(state_toclean, reverse = TRUE)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Erika Tyagi, Chase Hommeyer, Grace DiLaura,  and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 

## Load inputs

Input files: 

* Utilities script
* Historical data

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code", "historical-data")
data_path <- file.path(base_path, "data", "inputs")
##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```

Load in all the historical data for this state!

```{r load data, message=FALSE}
df <- load_data(data_path, 
                "11420",
                filter_state = state_toclean) 
df_typed <- type_convert(df)
```

Fill in missing values of `Staff.Deaths` with non-missing values of `Staff.Death`, when those exist. Do the same with `Resident.Death`/`Resident.Deaths`. 

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) %>% # rm 100% missing cols 
  mutate(Zipcode = as.character(Zipcode),
         County.FIPS = as.character(County.FIPS))

# if there are any other similarly-named columns in the results below, pay attention!
df_out %>%
  select(contains("Staff")) %>% 
  names()

df_out %>%
  select(contains("Resident")) %>% 
  names()
```

```{r}
df_out %<>% 
  mutate(Resident.Deaths = behindbarstools::coalesce_with_warnings(Resident.Deaths, Resident.Death)) 
```

```{r create date var}
df_out <- df_out %>%
  mutate(Date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r standardize facility names}
df_mid <- behindbarstools::clean_facility_name(df_out, debug = TRUE) 
```

```{r filter out federal facilities}
df_mid <- df_mid %>%
  filter(federal_bool == FALSE)
```

```{r facility merge checks, include = FALSE}
# show instances where merge didn't identify a clean name
# check if it's in fac_data
df_mid %>%
  filter(name_match == FALSE) %>% 
  select(scrape_name_clean) %>%
  unique()
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_mid, Date, Name))
see_if(nrow(df_mid) == nrow(distinct(df_mid, Name, Date)))

# if there are two values, sum them 
df_comb <- df_mid %>% 
  behindbarstools::group_by_coalesce(., Name, Facility.ID, Date, 
                                     .ignore = "scrape_name_clean",
                                     .method = "sum") 

assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, Date)))
```

Filter down and re-order columns in order to row bind them to latest data.

```{r}
df_hist <- behindbarstools::reorder_cols(df_comb)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Residents.Deaths = Resident.Deaths) %>%
  select(-c(Website, Resident.Deaths, Resident.Death, Count.ID, Facility)) 

df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE) %>%
  select_if(~sum(!is.na(.)) > 0) # rm all NA

```

Change anomalous state-wide cumulative infection counts to NA.

```{r}

df_hist_final_edit <- df_hist_final %>%
  mutate(Residents.Deaths = ifelse((Name == "CUMMINS UNIT PRISON" & Date == as.Date("2020-09-18")),
                                      NA,
                                      Residents.Deaths),
          Residents.Deaths = ifelse((Name == "CUMMINS UNIT PRISON" & Date == as.Date("2020-09-23")),
                                      NA,
                                      Residents.Deaths),
         Staff.Confirmed = ifelse((Name == "STATEWIDE" & Date > as.Date("2020-10-05")),
                            NA,
                            Staff.Confirmed),
         ) %>%
  filter(!(Date == as.Date("2020-09-20") & Name == "STATEWIDE")) %>%
  ## NB: before 4/28, our data does not line up with raw sources for staff.confirmed (seems to actually be staff.active) 
  select(-Staff.Confirmed)
  
  
```

Prep pre-November data for writing to the server. 
```{r}
df_hist_towrite <- prep_server_data(df_hist_final_edit, state_abbrev)

skim(df_hist_towrite) # double-check everything in the data 
```

Write pre-November data to the server. 

```{r}
srvr_outfile_name <- glue('1-pre-november-{state_abbrev}.csv')

write_csv(df_hist_towrite, file.path(base_path, "data", "pre-nov", srvr_outfile_name))

sync_remote_files(srvr_outfile_name) # ignore this line if you're not hope!
```

