---
title: "Clean AK - COVID-19 Behind Bars Historical Data Cleaning"
author: "Meichen Chen"
date: "4/16/21"
output: html_document
---

```{r package setup, include=FALSE}
rm(list=ls())
##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr", "skimr", "plotly")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
options(skimr_strip_metadata = FALSE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
library(behindbarstools)
help(package=behindbarstools)

##Define state
state_toclean <- "Alaska"
state_abbrev <- behindbarstools::translate_state(state_toclean, reverse = TRUE)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Erika Tyagi, Chase Hommeyer, Grace DiLaura,  and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 

## Load inputs

Input files: 

* Utilities script
* Historical data

```{r load inputs, echo=FALSE}
#
base_path <- file.path("~", "UCLA", "code", "historical-data")
data_path <- file.path(base_path, "data", "inputs")
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())

##Load utilities function
# source("~/Desktop/historical-data/R/0-utilities.R", local = knitr::knit_global())
```

Load in all the historical data for this state!

```{r load data, message=FALSE}
df <- load_data(data_path,
               "11420",
               filter_state = state_toclean)

# keep all data in a folder called Massachusetts
# df <- load_data("~/Desktop/Behind Bars/Massachusetts", "11420",
#                 filter_state = state_toclean) 
df_typed <- type_convert(df)
```

Fill in missing values of `Staff.Deaths` with non-missing values of `Staff.Death`, when those exist. Do the same with `Resident.Death`/`Resident.Deaths`. 

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) %>% # rm 100% missing cols 
  mutate(Zipcode = as.character(Zipcode),
         County.FIPS = as.character(County.FIPS))

# if there are any other similarly-named columns in the results below, pay attention!
df_out %>%
  select(contains("Staff")) %>% 
  names()

df_out %>%
  select(contains("Resident")) %>% 
  names()
```

```{r}

df_out %<>% 
  mutate(Staff.Deaths = behindbarstools::coalesce_with_warnings(Staff.Deaths, Staff.Death),
         Resident.Deaths = behindbarstools::coalesce_with_warnings(Resident.Deaths, Resident.Death)) 
```

```{r create date var}
df_out <- df_out %>%
  mutate(Date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r standardize facility names}
df_mid <- behindbarstools::clean_facility_name(df_out, debug = TRUE) 
```

```{r filter out federal facilities}
df_mid <- df_mid %>%
  filter(federal_bool == FALSE)
```

Show instances where the  didn't identify a clean name. 

- Stop and add these to facility spellings and/or facility data, here: https://docs.google.com/spreadsheets/d/1tAhD-brnuqw0s55QXM-xYLPsyD-rNrqHbAVIbxSOMwI/edit#gid=0

- Reference the current working versions of both sheets, available here: https://github.com/uclalawcovid19behindbars/facility_data/tree/master/data 

```{r facility name cleaning checks}
# check if it's in fac_data
df_mid %>%
  filter(name_match == FALSE) %>% 
  select(scrape_name_clean) %>%
  unique()
```

Remove rows when we didn't get a name at all.
```{r}
df_mid <- df_mid %>%
  filter(!is.na(Name))
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_mid, Date, Name))
see_if(nrow(df_mid) == nrow(distinct(df_mid, Name, Date)))

# if there are two values, sum them 
df_comb <- df_mid %>% 
  behindbarstools::group_by_coalesce(., Name, Facility.ID, Date, 
                                     .ignore = "scrape_name_clean",
                                     .method = "sum") 

assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, Date)))
```

Assertions / sanity checks ! 
```{r}
# make sure there's only 1 clean name per facility ID! 
df_comb %>% 
  group_by(Facility.ID) %>% 
  summarise(n_name = n_distinct(Name)) %>% 
  filter(n_name > 1)

# make sure there's only 1 facility ID per clean name!
df_comb %>% 
    group_by(Name) %>% 
    summarise(n_fac_ID = n_distinct(Facility.ID)) %>% 
    filter(n_fac_ID > 1)
```

Filter down and re-order columns in order to row bind them to latest data.

```{r}
df_hist <- behindbarstools::reorder_cols(df_comb)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Residents.Deaths = Resident.Deaths) %>%
  select(-c(Website, Resident.Deaths, Resident.Death, Count.ID, Facility)) 

df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE) %>%
  select_if(~sum(!is.na(.)) > 0) # rm all NA
```

Make plots for the data, and take a look at any weirdness. 

```{r plot cases/deaths}
res_conf_stacked <- df_hist_final %>% 
  mutate(Facility.ID = as.character(Facility.ID)) %>%
  arrange(Facility.ID) %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = Residents.Confirmed, 
             group = Facility.ID)) +
  theme(legend.position = "none") + 
  geom_line(aes(colour = Name), position = "stack", alpha = 0.4, size = 0.5) + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y") 
ggplotly(res_conf_stacked)

res_death_stacked <- df_hist_final %>% 
  mutate(Facility.ID = as.character(Facility.ID)) %>%
  arrange(Facility.ID) %>%
  ggplot(data = ., 
         aes(x = Date, 
             y = Residents.Deaths, 
             group = Facility.ID)) +
  theme(legend.position = "none") + 
  geom_line(aes(colour = Name), position = "stack", alpha = 0.4, size = 0.5) + 
  scale_x_date(date_breaks = "2 months", date_labels = "%m/%y") 
ggplotly(res_death_stacked)


# create a testing file 
testing_cases <- flag_noncumulative_cases(df_hist_final, Name)
testing_deaths <- flag_noncumulative_deaths(df_hist_final, Name, Residents.Deaths)

# lag cases overall 
all_cases <- testing_cases %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_cases, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
         scale_x_date(date_minor_breaks = "1 month", date_labels = "%m/%y", 
                      date_breaks = "1 month") + 
    labs(x = "Date",
      y = "lag_change_cases")

ggplotly(all_cases)

# lag deaths overall
all_deaths <- testing_deaths %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_deaths, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
    scale_x_date(date_labels = "%m") + 
    labs(x = "Date",
      y = "lag_change_deaths")
ggplotly(all_deaths)

```

Make facility-specific plots, and take a look at any weirdness.

```{r}
# will throw a warning if path already exists

plot_path <- file.path(base_path, "plots", state_abbrev)
dir.create(file.path(plot_path))
dir.create(file.path(plot_path, "cases"))
dir.create(file.path(plot_path, "deaths"))
# base_path <- "~/Desktop/historical-data"
# plot_path <- file.path("~/Desktop/historical-data", "plots", state_abbrev)
# dir.create(file.path(plot_path))
# dir.create(file.path(plot_path, "cases"))
# dir.create(file.path(plot_path, "deaths"))

# plot lag counts by facility
lag_case_plots <- plot_lags(testing_cases, "Date", 
                            y_var = "lag_change_cases",
                            grp_var = Name)
# lag_case_plots$plot # view them by un-commenting this

## save lag case plots
for (i in 1:nrow(lag_case_plots)){
  facility_name <- lag_case_plots$Name[[i]]
  ggsave(paste0(facility_name, "_LagChangeCases.png"), lag_case_plots$plot[[i]],
         path = file.path(base_path, "plots", state_abbrev, "cases"))
}

lag_death_plots <- plot_lags(testing_deaths, "Date", 
                            y_var = "lag_change_deaths",
                            grp_var = Name)
# save lag death plots
for (i in 1:nrow(lag_death_plots)){
  facility_name <- lag_death_plots$Name[[i]]
  ggsave(paste0(facility_name, "_LagChangeDeaths.png"), lag_death_plots$plot[[i]],
         path = file.path(base_path, "plots", state_abbrev, "deaths"))
}
```
Change anomalous state-wide cumulative infection counts that were not confirmed in the WayBack Machine. 

```{r}
df_hist_final_edit <- df_hist_final %>%
  mutate(Residents.Confirmed = ifelse((Name == "STATEWIDE" & Date == as.Date("2020-10-01")),
                                      NA,
                                      Residents.Confirmed),
         Residents.Confirmed = ifelse((Name == "STATEWIDE" & Date == as.Date("2020-09-04")),
                                      NA,
                                      Residents.Confirmed),
        Residents.Confirmed = ifelse((Name == "STATEWIDE" & Date == as.Date("2020-11-04")),
                            240,
                            Residents.Confirmed),
        Residents.Tested = ifelse((Name == "STATEWIDE" & Date == as.Date("2020-09-04")),
                            NA,
                            Residents.Tested),
        Residents.Tested = ifelse(Date == as.Date("2020-09-04") | Date == as.Date("2020-10-01"),
                    NA,
                    Residents.Tested)
         ) %>%
  select(-Staff.Confirmed,    ## from a news article, only one facility for five months 
         -Staff.Deaths)       ## ambiguous source
```


Find date spans / week spans with no data. In instances where the count went down by one, it could be that a PDF was misread. 

```{r}
dates <- df_hist_final_edit %>%
  arrange(Date) %>%
  count(Date)
dates

ggplot(data = dates, 
       aes(x = Date, y = n)) +
  geom_bar(stat="identity") +
  labs(x = "Date",
    y = "n instances")
```

Marshall Project / AP data comparison.

```{r}
mpap <- read_mpap_data(all_dates = TRUE) %>%
  filter(State == state_toclean)

ourdat_tocompare <- df_hist_final_edit %>% 
  filter(Jurisdiction == "state") %>%
  group_by(Date) %>%
  summarise(Residents.Confirmed = sum_na_rm(Residents.Confirmed),
            Residents.Deaths = sum_na_rm(Residents.Deaths),
            # Staff.Confirmed = sum_na_rm(Staff.Confirmed)
          # Residents.Recovered = sum_na_rm(Residents.Recovered),
           #Staff.Recovered = sum_na_rm(Staff.Recovered)
)
  
comparison <- ourdat_tocompare %>%
  full_join(mpap, by = "Date", suffix = c("_cbb", "_mpap")) %>%
  arrange(Date) %>%
  select(-State) %>% 
  select(Date, 
         starts_with("Residents.Confirmed"), 
         starts_with("Residents.Deaths"),
      #   starts_with("Residents.Recovered"),
       # starts_with("Staff.Confirmed")
)
```

```{r}
## looking for statewide drop on 2020-10-21, 2020-11-05
df_hist_final_edit %>% 
  filter(Name == "STATEWIDE") %>% 
  select(Name, Date, Residents.Confirmed) %>%
  View()

## come back and check this in 
ak_checking <- read_scrape_data(all_dates = TRUE, state = "Alaska")
ak_checking %>% 
  filter(Name == "STATEWIDE") %>%
  select(Name, Date, Residents.Confirmed) %>%
  View()
```


Prep pre-November data for writing to the server. 

```{r}
df_hist_towrite <- prep_server_data(df_hist_final_edit, state_abbrev)

skim(df_hist_towrite) # double-check everything in the data 
```

Write pre-November data to the server. 

```{r}
srvr_outfile_name <- glue('1-pre-november-{state_abbrev}.csv')

write_csv(df_hist_towrite, file.path(base_path, "data", "pre-nov", srvr_outfile_name))

sync_remote_files(srvr_outfile_name) # ignore this line if you're not hope!
```
