---
title: "Clean North Carolina - COVID-19 Behind Bars Historical Data Cleaning"
author: "Hope Johnson"
date: "12/07/2020"
output: html_document
---

```{r package setup, include=FALSE}
##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
help(package=behindbarstools)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Chase Hommeyer, Grace DiLaura, and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 


## Load inputs

Input files: 

* Utilities script
* Historical data

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code", "historical", "historical-data")
data_path <- file.path(base_path, "data", "inputs")
##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```


```{r load data, message=FALSE}
df <- load_data(data_path, 
                "11420",
                filter_state = "Wisconsin") 
df_typed <- type_convert(df)
```

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) # rm 100% missing cols 
names(df_out) 
```

```{r create date var}
df_out <- df_out %>%
  mutate(date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r standardize facility names}
df_mid <- behindbarstools::clean_facility_name(df_out) %>%
  select(-name_xwalk)
```

```{r filter out federal facilities}
df_mid <- df_mid %>%
  filter(federal_bool == FALSE)
```

```{r facility merge checks, include = FALSE}
# show instances where merge didn't identify a clean name:
## DIVISION OF JUVENILE CORRECTIONS FACILITIES is missing
## DIVISION OF JUVENILE CORRECTIONS FIELD is missing
df_mid %>%
  filter(Name == "NULL") %>% 
  select(scrape_name_clean)
```

Remove observations where we didn't pick up a clean name.

```{r drop null name observations}
df_mid <- df_mid %>%
  filter(!is.na(scrape_name_clean),
         Name != "NULL")
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_mid, date, Name))
see_if(nrow(df_mid) == nrow(distinct(df_mid, Name, date)))

df_comb <- df_mid %>% 
  behindbarstools::group_by_coalesce(., Name, date, 
                                     .ignore = c("scrape_name_clean", "Facility", "source"))

assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, date)))
```

Filter down and re-order columns in order to row bind them to latest data.

```{r}
df_hist <- behindbarstools::reorder_cols(df_comb)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Residents.Deaths = Resident.Deaths,
         Date = date,
         jurisdiction = Facility,
         ID = Count.ID) %>%
  select(-c(Website, Resident.Deaths, date)) 
df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE)
## add back in Residents.Tested (this will only be present for historical data)
df_hist_final$Residents.Tested <- df_hist$Residents.Tested
```

Add in more recently scraped data (from November until present). 

First, read it in and clean it. 

```{r add in recently scraped data}
slist <- read_rds(file.path(data_path, "scrape_details.rds"))
hist_dat <- slist[["hist_data"]]
today_dat <- slist[["current_data"]]
recent_dat <- hist_dat %>%
  bind_rows(today_dat)

recent_dat <- recent_dat %>%
  mutate(Date = lubridate::mdy(Date)) %>%
  filter(State == "Wisconsin",
         jurisdiction != "federal",
         Date > as.Date('2020-11-04')) 

## TO DO : add NEW LISBON CORRECTIONAL INSTITUTIO into xwalk, and: 
## RACINE CORRECTIONAL INSTITUTION 

## TO DO: Come back to this once I've resolved "auxilary", "youth facilities", etc

recent_dat_cln <- behindbarstools::clean_facility_name(recent_dat) %>%
  select(-name_xwalk) 

recent_dat_cln %>%
  filter(Name == "NULL") %>%
  View()

recent_dat_cln <- recent_dat_cln %>%
  filter(Name != "NULL")

recent_comb <- recent_dat_cln %>%
  behindbarstools::group_by_coalesce(., Name, Date,
                                     .ignore = c("source", "scrape_name_clean", "jurisdiction")) 

recent_final <- behindbarstools::reorder_cols(recent_comb, rm_extra_cols = TRUE)
recent_final$Residents.Tested <- NA
```
Then, bind it to the historical cleaned data from this script. 

```{r bind historical and recent}
all_dat <- bind_rows(df_hist_final, recent_final)
n_distinct(all_dat$Name)
```
Merge in facility information (address, city, county, etc.).

TO DO : Fill in missing info, especially lat/lon, in "facility_info" sheet.

```{r}
fac_info <- behindbarstools::read_fac_info() %>%
  filter(State == "Wisconsin") %>%
  mutate(Name = behindbarstools::clean_fac_col_txt(Name))
final_dat <- left_join(all_dat, fac_info, 
                  by = "Name",
                  suffix = c(".x", ""))

final_dat %<>% 
  mutate(Latitude = behindbarstools::coalesce_with_warnings(Latitude.x, Latitude),
         Longitude = behindbarstools::coalesce_with_warnings(Longitude.x, Longitude)) %>%
  select(-ends_with(".x"))
```

Find non-cumulative counts, separate those into "active" and "confirmed". Confirmed = active + recovered + deaths.

```{r plot cases/deaths}
test <- flag_noncumulative_cases(final_dat, Name)

# lag cases overall 
test %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_cases, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
         scale_x_date(date_minor_breaks = "1 month", date_labels = "%m/%y", 
                      date_breaks = "1 month") + 
    labs(x = "Date",
      y = "lag_change_cases")

# plot lag counts by facility
lag_case_plots <- plot_lags(test, "Date", 
                            y_var = "lag_change_cases",
                            grp_var = Name)
## save lag case plots
# for (i in 1:nrow(lag_case_plots)){
#   facility_name <- lag_case_plots$Name[[i]]
#   ggsave(paste0(facility_name, "_LagChangeCases.png"), lag_case_plots$plot[[i]],
#          path = file.path(base_path, "plots", "WI"))
# }
```

Find date spans / week spans with no data. In instances where the count went down by one, it could be that a PDF was misread. 

```{r}
dates <- test %>%
  arrange(Date) %>%
  count(Date)
dates

ggplot(data = dates, 
       aes(x = Date, y = n)) +
  geom_bar(stat="identity") +
  labs(x = "Date",
    y = "n instances")
```

```{r find infection spikes, include=FALSE }
## Find historical spikes over time.
## get 7-day percentage change in cases
percentage_change <- final_dat %>%
    dplyr::arrange(Name, Date) %>%
    mutate(
        week_ago = dplyr::lag(Residents.Confirmed, n = 7),
        seven_day_change = (Residents.Confirmed - week_ago) / (week_ago + 0.00001) ,
        n_increase = Residents.Confirmed - week_ago,
        denom_zero = ifelse(week_ago == 0, TRUE, FALSE)
    ) %>%
    select(State, Name, Date, Residents.Confirmed, 
           week_ago, seven_day_change, denom_zero, n_increase) %>%
    ungroup()

# View(percentage_change)
```

Finally, merge in population data. This isn't working right now because of naming issues!

TO DO: also check "/Users/hope/Downloads/wi.csv"

````{r merge-population}
pop <- read_csv('https://raw.githubusercontent.com/uclalawcovid19behindbars/Population/main/Merg_Pop.csv') %>%
  mutate(Name = behindbarstools::clean_fac_col_txt(Name, to_upper = TRUE)) %>%
  filter(State == "Wisconsin")
head(pop)

pop_alt <- read_csv(file.path("~", "Downloads", "wi.csv")) %>%
  mutate(Name_cln = behindbarstools::clean_fac_col_txt(Name, to_upper = TRUE)) %>%
  mutate(population_date = Date ) %>%
  select(-c(State, Date))

dat_with_pop <- test %>%
  left_join(pop_alt, by = c("Name" = "Name_cln",
                            "Date" = "population_date"))

table(is.na(dat_with_pop$Population))
```


```{r write csv}
write_csv(final_dat, file.path(base_path, "data", "wi_adult_facility_covid_counts_historical.csv"))
```
