---
title: "Clean North Carolina - COVID-19 Behind Bars Historical Data Cleaning"
author: "Hope Johnson"
date: "12/21/2020"
output: html_document
---

```{r package setup, include=FALSE}
##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
help(package=behindbarstools)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Erika Tyagi, Chase Hommeyer, Grace DiLaura,  and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 


## Load inputs

Input files: 

* Utilities script
* Historical data

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code", "historical", "historical-data")
data_path <- file.path(base_path, "data", "inputs")
##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```


```{r load data, message=FALSE}
df <- load_data(data_path, 
                "11420",
                filter_state = "North Carolina") 
df_typed <- type_convert(df)
```

Fill in missing values of `Staff.Deaths` with non-missing values of `Staff.Death`, when those exist. Do the same with `Resident.Death`/`Resident.Deaths`. 

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) # rm 100% missing cols 
names(df_out) # stopped here.

df_out %<>% 
  mutate(Resident.Deaths = behindbarstools::coalesce_with_warnings(Resident.Deaths, Resident.Death)) 
```

```{r create date var}
df_out <- df_out %>%
  mutate(date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r standardize facility names}
df_mid <- behindbarstools::clean_facility_name(df_out, debug = TRUE) 
```

```{r filter out federal facilities}
df_mid <- df_mid %>%
  filter(federal_bool == FALSE)
```

```{r facility merge checks, include = FALSE}
# show instances where merge didn't identify a clean name
df_mid %>%
  filter(name_match == FALSE) %>% 
  select(scrape_name_clean)
```

Remove observations where we didn't pick up a clean name.

```{r drop null name observations}
df_mid <- df_mid %>%
  filter(!is.na(scrape_name_clean),
         name_match != FALSE)
```

Remove observation where we clearly mis-scraped the data.

```{r drop mis-scrape}
df_mid <- df_mid %>%
  filter(date != "2020-08-29")
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_mid, date, Name))
see_if(nrow(df_mid) == nrow(distinct(df_mid, Name, date)))

# if there are two values, sum them 
df_comb <- df_mid %>% 
  behindbarstools::group_by_coalesce(., Name, date, 
                                     .ignore = "scrape_name_clean",
                                     .method = "sum")

assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, date)))
```

Filter down and re-order columns in order to row bind them to latest data.

```{r}
df_hist <- behindbarstools::reorder_cols(df_comb)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Residents.Deaths = Resident.Deaths,
         Date = date,
         jurisdiction = Facility,
         ID = Count.ID) %>%
  select(-c(Website, Resident.Deaths, Resident.Death, date)) 
df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE)
## add back in Residents.Tested (this will only be present for historical data)
df_hist_final$Residents.Tested <- df_hist$Residents.Tested
```

Add in more recently scraped data (from November until present). 

First, read it in and clean it. 

```{r add in recently scraped data}
recent_dat <- behindbarstools::read_scrape_data(all_dates = TRUE, 
                                                state = "North Carolina", 
                                                debug = TRUE) %>%
  filter(jurisdiction != "federal",
       Date > as.Date('2020-11-04')) 

## view facility names without a match in the xwalk
recent_dat %>%
  filter(name_match == "FALSE") %>% 
  select(scrape_name_clean, Name)

recent_dat_cln <- recent_dat %>%
  filter(name_match != FALSE) %>%
  select(-xwalk_name_clean)

recent_final <- behindbarstools::reorder_cols(recent_dat_cln, rm_extra_cols = TRUE)
recent_final$Residents.Tested <- NA
```
Then, bind it to the historical cleaned data from this script. 

```{r bind historical and recent}
all_dat <- bind_rows(df_hist_final, recent_final)
n_distinct(all_dat$Name)
```
Merge in facility information (address, city, county, etc.).

```{r}
fac_info <- behindbarstools::read_fac_info() %>%
  filter(State == "North Carolina") %>%
  mutate(Name = behindbarstools::clean_fac_col_txt(Name))
final_dat <- left_join(all_dat, fac_info, 
                  by = "Name",
                  suffix = c(".x", "")) %>%
  select(-ends_with(".x"))
```

Find non-cumulative counts, separate those into "active" and "confirmed". Confirmed = active + recovered + deaths.

```{r plot cases/deaths}
test_nc <- flag_noncumulative_cases(final_dat, Name)
test_nc <- flag_noncumulative_deaths(test_nc, Name, Residents.Deaths)

# lag cases overall 
test_nc %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_cases, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
         scale_x_date(date_minor_breaks = "1 month", date_labels = "%m/%y", 
                      date_breaks = "1 month") + 
    labs(x = "Date",
      y = "lag_change_cases")

# lag deaths overall
test_nc %>%
  filter(Name != "STATEWIDE") %>%
  ggplot(data = ., 
         aes(x = Date, y = lag_change_deaths, group = Name)) +
    geom_line(alpha=0.6 , size=.5, color = "black") +
    scale_x_date(date_labels = "%m") + 
    labs(x = "Date",
      y = "lag_change_deaths")

# plot lag counts by facility
lag_case_plots <- plot_lags(test_nc, "Date", 
                            y_var = "lag_change_cases",
                            grp_var = Name)
lag_case_plots$plot

# save lag case plots
# for (i in 1:nrow(lag_case_plots)){
#   facility_name <- lag_case_plots$Name[[i]]
#   ggsave(paste0(facility_name, "_LagChangeCases.png"), lag_case_plots$plot[[i]],
#          path = file.path(base_path, "plots", "NC", "cases"))
# }

## confirmed Wilkes CC via way back machine
## confirmed Marion CI via way back machine
## confirmed Foothills CI via way back machine
## confirmed Central Prison via way back machine
## confirmed Alexander CI via way back machine

lag_death_plots <- plot_lags(test_nc, "Date", 
                            y_var = "lag_change_deaths",
                            grp_var = Name)
# save lag death plots
# for (i in 1:nrow(lag_death_plots)){
#   facility_name <- lag_death_plots$Name[[i]]
#   ggsave(paste0(facility_name, "_LagChangeDeaths.png"), lag_death_plots$plot[[i]],
#          path = file.path(base_path, "plots", "NC", "deaths"))
# }
```

Find date spans / week spans with no data. In instances where the count went down by one, it could be that a PDF was misread. 

```{r}
dates <- final_dat %>%
  arrange(Date) %>%
  count(Date)
dates

ggplot(data = dates, 
       aes(x = Date, y = n)) +
  geom_bar(stat="identity") +
  labs(x = "Date",
    y = "n instances")
```

```{r find infection spikes, include=FALSE }
## Find historical spikes over time.
## get 7-day percentage change in cases
percentage_change <- final_dat %>%
    dplyr::arrange(Name, Date) %>%
    mutate(
        week_ago = dplyr::lag(Residents.Confirmed, n = 7),
        seven_day_change = (Residents.Confirmed - week_ago) / (week_ago + 0.00001) ,
        n_increase = Residents.Confirmed - week_ago,
        denom_zero = ifelse(week_ago == 0, TRUE, FALSE)
    ) %>%
    select(State, Name, Date, Residents.Confirmed, 
           week_ago, seven_day_change, denom_zero, n_increase) %>%
    ungroup()

# View(percentage_change)
```

Finally, merge in population data. This isn't working right now because of naming issues!

````{r merge-population}
pop <- read_csv('https://raw.githubusercontent.com/uclalawcovid19behindbars/Population/main/initial/Merg_Pop.csv') %>%
  mutate(Name = behindbarstools::clean_fac_col_txt(Name, to_upper = TRUE)) %>%
  filter(State == "North Carolina")
head(pop)

dat_with_pop <- final_pop %>%
  left_join(pop, by = "Name")

table(is.na(dat_with_pop$Population))
```


```{r write csv}
out <- final_dat %>%
  rename(hifld_pop = POPULATION) %>%
  reorder_historical_cols(rm_extra_cols = TRUE)

write_csv(out, file.path(base_path, "data", "NC_adult_facility_covid_counts_historical.csv"))
```
