---
title: "Clean Wisconsin - COVID-19 Behind Bars Historical Data Cleaning"
author: "Hope Johnson"
date: "2/5/21"
output: html_document
---

```{r package setup, include=FALSE}
rm(list=ls())

##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
help(package=behindbarstools)

##Define state
state_toclean <- "Wisconsin"
state_abbrev <- behindbarstools::translate_state(state_toclean, reverse = TRUE)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Chase Hommeyer, Grace DiLaura, and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 


## Load inputs

Input files: 

* Utilities script
* Historical data

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code", "historical-data")
data_path <- file.path(base_path, "data", "inputs")
##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```


```{r load data, message=FALSE}
df <- load_data(data_path, 
                "11420",
                filter_state = "Wisconsin") 
df_typed <- type_convert(df)
```

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) # rm 100% missing cols 
names(df_out) 
```

```{r create date var}
df_out <- df_out %>%
  mutate(date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r standardize facility names}
df_mid <- behindbarstools::clean_facility_name(df_out, debug = TRUE)
```

```{r filter out federal facilities}
df_mid <- df_mid %>%
  filter(federal_bool == FALSE)
```

```{r facility merge checks, include = FALSE}
df_mid %>%
  filter(name_match == "FALSE") %>% 
  select(scrape_name_clean) %>%
  unique()
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_mid, date, Name))
see_if(nrow(df_mid) == nrow(distinct(df_mid, Name, date)))

df_comb <- df_mid %>% 
  behindbarstools::group_by_coalesce(., Name, date, 
                                     .ignore = c("scrape_name_clean", "Facility", "source"))

assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, date)))
```

Filter down and re-order columns in order to row bind them to latest data.

```{r}
df_hist <- behindbarstools::reorder_cols(df_comb)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Date = date) %>%
  select(-c(Website, date, Count.ID, Facility)) 

df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE) %>%
  select_if(~sum(!is.na(.)) > 0) 
```

Prep pre-November data for writing to the server. 
```{r}
df_hist_towrite <- prep_server_data(df_hist_final, state_abbrev)

skim(df_hist_towrite) # double-check everything in the data 
```

Write pre-November data to the server. 

```{r}
srvr_outfile_name <- glue('1-pre-november-{state_abbrev}.csv')

write_csv(df_hist_towrite, file.path(base_path, "data", "pre-nov", srvr_outfile_name))

sync_remote_files(srvr_outfile_name) # ignore this line if you're not hope!
```