---
title: "Clean Pennsylvania - COVID-19 Behind Bars Historical Data Cleaning"
author: "Hope Johnson"
date: "2/5/2021"
output: html_document
---

```{r package setup, include=FALSE}
##Define package list
Packages<-c("tidyverse", "glue", "assertthat", "stringr", "lubridate",
            "devtools", "magrittr", "skimr")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
devtools::install_github("uclalawcovid19behindbars/behindbarstools")
help(package=behindbarstools)
```

## Intro & Credits

This script is used to clean one state in the historical data concerning COVID-19 in state and federal prisons. Contributors to the historical data cleaning efforts include Hope Johnson, Michael Everett, Neal Marquez, Erika Tyagi, Chase Hommeyer, Grace DiLaura, and Kalind Parish. Contributors to larger project include Sharon Dolovich, Aaron Littman, Danielle Flores, Poornima Rajeshwar, Victoria Rossi, and many others. 


## Load inputs

Input files: 

* Utilities script
* Historical data
* Date range to clean 
* Facility name look-up table 

```{r load inputs, echo=FALSE}
base_path <- file.path("~", "UCLA", "code", "historical", "historical-data")
data_path <- file.path(base_path, "data", "inputs")
##Load utilities function
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```


```{r load data, message=FALSE}
df <- load_data(data_path, 
                "11420",
                filter_state = "Pennsylvania") 
df_typed <- type_convert(df)
```

Fill in missing values of `Staff.Deaths` with non-missing values of `Staff.Death`, when those exist. Do the same with `Resident.Death`/`Resident.Deaths`. 

```{r initial cleaning}
df_out <- df_typed %>%
  select(!starts_with("...")) %>%
  select(!starts_with("lots")) %>%
  select(!starts_with("Bad")) %>%
  select(!c("V2", "V4", "V5", "V7", "V8", "V10")) %>%
  select(!c("Facility.", "Coder", "Housing.Type")) %>%
  select_if(~sum(!is.na(.)) > 0) # rm 100% missing cols 

df_out %<>% 
  mutate(Resident.Deaths = behindbarstools::coalesce_with_warnings(Resident.Deaths, Resident.Death),
         Staff.Deaths = behindbarstools::coalesce_with_warnings(Staff.Deaths, Staff.Death)) %>% 
  mutate(Residents.Confirmed = na_if(Residents.Confirmed, "n/a")) %>%
  mutate(Residents.Confirmed = as.numeric(Residents.Confirmed)) %>%
  select(-Resident.Death,
         -Resident.Deaths,
         -Staff.Death) # only keep Staff.Deaths and Residents.Deaths
```

```{r create date var}
df_out <- df_out %>%
  mutate(date = as_date(sheet_name, format = "%Om.%d.%y"))
```

```{r standardize facility names}
df_mid <- behindbarstools::clean_facility_name(df_out, debug = TRUE) 
```

```{r filter out federal facilities}
df_mid <- df_mid %>%
  filter(federal_bool == FALSE)
```

```{r facility merge checks, include = FALSE}
# show instances where merge didn't identify a clean name
df_mid %>%
  filter(name_match == FALSE) %>% 
  select(scrape_name_clean) %>%
  unique()
```

Figure out duplicate date/facilities, concatenate those instances from multiple rows into one. This most often occurs because we scraped death data and infections data from separate tables.

```{r concat duplicate date/facilities}
nrow(distinct(df_mid, date, Name))
see_if(nrow(df_mid) == nrow(distinct(df_mid, Name, date)))

# if there are two values, sum them 
df_comb <- df_mid %>% 
  behindbarstools::group_by_coalesce(., Name, date, Facility.ID,
                                     .ignore = "scrape_name_clean",
                                     .method = "sum")

assert_that(nrow(df_comb) == nrow(distinct(df_comb, Name, date)))
```

```{r}
df_hist <- behindbarstools::reorder_cols(df_comb)
df_hist_final <- df_hist %>%
  mutate(source = Website,
         Date = date) %>%
  select(-c(Website, date, Count.ID, Facility)) 

df_hist_final <- behindbarstools::reorder_cols(df_hist_final, rm_extra_cols = TRUE) %>%
  select_if(~sum(!is.na(.)) > 0) %>% # get rid of all-NA columns for easier row-binding
  mutate(Zipcode = as.character(Zipcode),
         County.FIPS = as.character(County.FIPS))

```

Prep pre-November data for writing to the server. 
```{r}
df_hist_towrite <- prep_server_data(df_hist_final, state_abbrev)

skim(df_hist_towrite) # double-check everything in the data 
```

Write pre-November data to the server. 

```{r}
srvr_outfile_name <- glue('1-pre-november-{state_abbrev}.csv')

write_csv(df_hist_towrite, file.path(base_path, "data", "pre-nov", srvr_outfile_name))

sync_remote_files(srvr_outfile_name) # ignore this line if you're not hope!
```