---
title: "Historical Data Monitoring"
author: "Hope Johnson, Neal Marquez, Michael Everett, Erika Tyagi"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Introduction

The purpose of this file is to generate a coherent report on the historical data from a given state. We aim to highlight case and death COVID-19 spikes within specific facilities and time periods. This document is a work-in-progress!

```{r load packages, warning=FALSE, message=FALSE, echo=FALSE}
##Define package list
library(behindbarstools)
Packages<-c("tidyverse", "glue", "knitr", "zoo")
.packages = Packages
##Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])
##Load packages into session 
lapply(.packages, require, character.only=TRUE)
##Load utilities function
base_path <- file.path("~", "UCLA", "code", "historical", "historical-data")
util_path <- file.path(base_path, "R", "0-utilities.R")
source(util_path, local = knitr::knit_global())
```
We are cleaning historical data, and not all of it is ready to monitor. The available states are:

```{r load historical data, warning=FALSE, message=FALSE, echo=FALSE}
data_path = file.path(base_path, "data")
## Available states for historical data monitoring
(clean_states = list.files(path = data_path, pattern = "*.csv"))

# eventually, cycle through these to create reports? 
state_to_clean = clean_states[4]

hist_data <- read_csv(file.path(data_path, state_to_clean))

# one of "historical population", "Residents.Population" (Feb 1st pop, or HIFLD when Feb. 1st unavailable)
best_pop_var <- sym("Population")
```

## Scraper Details

```{r warning=FALSE, message=FALSE, echo=FALSE}
state_to_clean <- unique(hist_data$State)

## Choose prison here for examples 
focus_prison <- "OSHKOSH CORRECTIONAL INSTITUTION"

prison_n <- hist_data %>%
    filter(jurisdiction != "Jail",
           jurisdiction != "county") %>%
    pull(Name) %>%
    unique() %>%
    length()

jail_n <- hist_data %>%
    filter(jurisdiction == "Jail" | 
           jurisdiction == "county") %>%
    pull(Name) %>%
    unique() %>%
    length()

fac_rep_deaths <- hist_data %>%
  filter_statewide_NA(Residents.Deaths) %>%
  group_by(Name) %>%
  summarize(cumulative_deaths = last(Residents.Deaths, order_by = Date))

fac_rep_cases <- hist_data %>%
  filter_statewide_NA(Residents.Confirmed) %>%
  group_by(Name) %>%
  summarize(cumulative_cases = last(Residents.Confirmed, order_by = Date)) 

fac_cases_deaths <- fac_rep_cases %>%
  full_join(fac_rep_deaths)

max_cases = max(fac_cases_deaths$cumulative_cases, na.rm = TRUE)
max_deaths = ifelse(is.infinite(max(fac_cases_deaths$cumulative_deaths, na.rm = TRUE)),
                    NA,
                    max(fac_cases_deaths$cumulative_deaths, na.rm = TRUE))
max_case_facility <- fac_cases_deaths$Name[fac_cases_deaths$cumulative_cases == max_cases]
max_deaths_facility <- ifelse(is.na(max_deaths),
                    NA,
                    fac_cases_deaths$Name[fac_cases_deaths$cumulative_deaths == max_deaths])
```

State: `r state_to_clean`

Date collected: **`r min(hist_data$Date)` - `r max(hist_data$Date)`**

Data collected from **`r prison_n`** prisons and **`r jail_n`** jails (ignoring Federal prisons contained within the state)

The facility in `r unique(hist_data$State)` with the highest number of cumulative cases over the course of the pandemic is `r max_case_facility`. This facility has had **`r max_cases`** cases total within the incarcerated population.  

The facility in `r unique(hist_data$State)` with the highest number of cumulative deaths over the course of the pandemic is `r max_deaths_facility`. This facility has had **`r max_deaths`** deaths total.  

```{r warning=FALSE, message=FALSE, echo=FALSE, results='asis'}
opts <- options(knitr.kable.NA = "")
fac_cases_deaths %>%
  head(n = 7) %>%
  arrange(-cumulative_cases) %>%
  kable(col.names = c('Facility Name', 'Cumulative cases (incarcerated pop.)', 'Cumulative deaths (incarcerated pop.)'))
```

## Testing Rates

We collect testing information when available. Up until ~November 2020, the testing numbers we collected could have been the total number of tests administered OR the total number of individuals tested. We did not keep track of which number was being reported by the state. After November 2020, we started collected exclusively the number of tests administered. 

```{r warning=FALSE, message=FALSE, echo=FALSE}
fac_t_admin_n <- hist_data %>%
  get_fac_n(Residents.Tadmin)

fac_res_tested <- hist_data %>%
  get_last_value(Residents.Tested)

fac_res_tested_n <- hist_data %>%
  get_fac_n(Residents.Tested)

fac_testing_rates <- hist_data %>%
  filter_statewide_NA(Residents.Tested) %>%
  mutate(res_tested_rate = Residents.Tested / !!best_pop_var,
         tadmin_rate = Residents.Tadmin / !!best_pop_var) %>%
  select(Date, Name, res_tested_rate, Residents.Tested, 
         tadmin_rate, Residents.Tadmin, !!best_pop_var) %>%
  arrange(res_tested_rate)

res_tested_plots <- plot_lags(fac_testing_rates, "Date", 
                            y_var = "res_tested_rate",
                            grp_var = Name,
                            y_lab = "Total testing rate"
                            ) %>%
  rename(res_tested_rate_plot = plot)

t_admin_plots <- plot_lags(fac_testing_rates, "Date",
                            y_var = "tadmin_rate",
                            grp_var = Name,
                            y_lab = "Tests administered rate"
                           ) %>%
  rename(tadmin_rate_plot = plot)

testing_plots <- res_tested_plots %>%
  left_join(t_admin_plots, by = "Name") 

## show plots if the data exists for any facilities
if(fac_res_tested_n > 0){ head(testing_plots$res_tested_rate_plot)}
if(fac_t_admin_n > 0){ head(testing_plots$tadmin_rate_plot)}
```

We have data on the number of **tests administered** for `r fac_t_admin_n` facilities. We have data on the number of **tests given** (ambiguous definition) for `r fac_res_tested_n` facilities.

## Cases by Facility

First, we calculate rolling averages to get the total average cases over the past 3 days, 7 days, and so-on. 

```{r rolling averages, warning=FALSE, message=FALSE, echo=FALSE}
## heavily relied on this article: https://www.storybench.org/how-to-calculate-a-rolling-average-in-r/ 

hist_data <- hist_data %>%
    dplyr::arrange(desc(Name)) %>% 
    dplyr::group_by(Name) %>% 
    dplyr::mutate(res.confirmed_03day = zoo::rollmean(Residents.Confirmed, k = 3, fill = NA),
                  res.confirmed_05day = zoo::rollmean(Residents.Confirmed, k = 5, fill = NA),
                  res.confirmed_07day = zoo::rollmean(Residents.Confirmed, k = 7, fill = NA),
                  res.confirmed_15day = zoo::rollmean(Residents.Confirmed, k = 15, fill = NA),
                  res.confirmed_21day = zoo::rollmean(Residents.Confirmed, k = 21, fill = NA)) %>% 
  dplyr::ungroup()

hist_data %>% 
  dplyr::arrange(desc(Date)) %>% 
  dplyr::filter(Name == focus_prison) %>% 
  dplyr::select(Name,
                Date,
                Residents.Confirmed,
                res.confirmed_03day:res.confirmed_07day) %>% 
  utils::head(5) %>%
  kable()

hist_data %>% 
  dplyr::filter(Name == focus_prison) %>% 
  tidyr::pivot_longer(names_to = "rolling_mean_key", 
                    values_to = "rolling_mean_value", 
                    cols = c(Residents.Confirmed, 
                             res.confirmed_07day,
                             res.confirmed_21day)) %>%
  ggplot2::ggplot(aes(x = Date, 
                      y = rolling_mean_value, 
                      color = rolling_mean_key)) +
  ggplot2::geom_line() +   
  ggplot2::labs(title = glue("One {state_to_clean} prison's rolling average total COVID infections"), 
                subtitle = focus_prison,
                y = "Infections", 
                color = "Metric",
                x = "Date") 

```

Next, we calculate new daily cases in each facility. 

```{r daily case counts, warning=FALSE, message=FALSE, echo=FALSE}
new_cases <- hist_data %>%
  group_by(Name, Date) %>% 
  dplyr::summarize(
    confirmed_sum = (sum(Residents.Confirmed, na.rm = TRUE))) %>% 
  mutate(new_confirmed_cases = confirmed_sum - dplyr::lag(x = confirmed_sum, n = 1, 
                                              order_by = Date)) %>% 
  dplyr::select(Name, 
                new_confirmed_cases, 
                Date) %>% 
  dplyr::left_join(., y = hist_data, 
                   by = c("Name", "Date")) %>% 
  dplyr::select(Name,
                State,
                Date,
                Residents.Confirmed,
                dplyr::contains("confirm"),
                best_pop_var)

new_cases %>% 
  dplyr::filter(Name == focus_prison) %>% 
  dplyr::select(Name, Date, Residents.Confirmed, new_confirmed_cases) %>% 
  utils::head() %>%
  kable()
```

Calculate rolling averages for new cases, and plot them by facility.

```{r rolling averages daily case counts, warning=FALSE, message=FALSE, echo=FALSE}
new_cases <- new_cases %>%
    dplyr::group_by(Name) %>% 
    dplyr::mutate(
      new_confirmed_03day = zoo::rollmean(new_confirmed_cases, k = 3, fill = NA),
      new_confirmed_05day = zoo::rollmean(new_confirmed_cases, k = 5, fill = NA),
      new_confirmed_07day = zoo::rollmean(new_confirmed_cases, k = 7, fill = NA),
      new_confirmed_15day = zoo::rollmean(new_confirmed_cases, k = 15, fill = NA),
      new_confirmed_21day = zoo::rollmean(new_confirmed_cases, k = 21, fill = NA)) %>% 
  dplyr::ungroup()

## pivot longer
new_case_tidy <- new_cases %>% 
  tidyr::pivot_longer(names_to = "new_conf_av_key", 
                    values_to = "new_conf_av_value", 
                    cols = c(new_confirmed_03day,
                             new_confirmed_07day)) %>% 
  dplyr::mutate(new_conf_av_key = dplyr::case_when(
    new_conf_av_key == "new_confirmed_03day" ~ "3-day new confirmed cases",
    new_conf_av_key == "new_confirmed_07day" ~ "7-day new confirmed cases",
    TRUE ~ NA_character_)) %>% 
  dplyr::select(Date, 
                State, 
                Name,
                new_conf_av_value, 
                new_conf_av_key)
```

```{r plot rolling averages, warning=FALSE, message=FALSE, echo=FALSE, out.height="500%" } 
min_date <- min(new_case_tidy$Date, na.rm = TRUE)
max_date <- max(new_case_tidy$Date, na.rm = TRUE)

## Daily rolling averages
## TO DO: some way of filtering these
new_cases %>% 
  filter(Name == focus_prison) %>%
    ggplot2::ggplot(aes(x = Date, 
                      y = new_confirmed_cases)) +
    geom_col(alpha = 3/10, linetype = 0) + 
    ggplot2::geom_line(data = (new_case_tidy %>% 
                         filter(Name == focus_prison)), 
                       mapping = aes(x = Date, 
                                     y = new_conf_av_value, 
                                     color = new_conf_av_key)) +  
    ggplot2::facet_wrap( ~ Name)  +
    ggplot2::labs(title = glue("{state_to_clean} rolling 3 and 7-day averages of new COVID cases"),
                  y = "New Cases",
                  color = "Metric:", 
                  x = "Date") + 
   scale_x_date(date_minor_breaks = "1 month", date_labels = "%m/%y", 
              date_breaks = "1 month") + 
      ggplot2::theme(legend.position = "top")
```

Up next, calculate by facility:

- Total cases 

- Total cases per capita per 100,000

- Daily new cases, 7-day rolling average per capita 

- Daily new cases, 7-day rolling average per capita per 100,000

```{r table, warning=FALSE, message=FALSE, echo=FALSE}
all_data <- hist_data %>%
  left_join(new_cases, by = c("Name", "Date"), 
            suffix = c("", ".x")) %>%
  select(-ends_with(".x")) %>%
  mutate(total_pc = Residents.Confirmed / Population, # cumulative per capita 
         total_pc_p100 = total_pc / 100000,           # cumulative per capita per 100,000
         daily_pc_7day = new_confirmed_07day / Population, # daily cases (seven day average) per capita
         daily_pc_7day_p100 = daily_pc_7day / 100000 ) # daily cases (seven day average) per capita per 100,000

all_data %>% 
  filter(total_pc > 0.2) %>%
  select(Name, Date, Population, total_pc, total_pc_p100,
         daily_pc_7day, daily_pc_7day_p100) %>%
  arrange(desc(total_pc_p100)) %>%
  kable(col.names = c('Facility Name', 'Date', 'Pop.',
                      'Cumulative cases per capita (incarcerated pop.)', 
                      'Per 100,000',
                      '7-day average, daily cases per capita (incarcerated pop.)', 
                      'Per 100,000'))
```

## Outlier Analysis

- Mahalanobis Distance ?

- Using IFR = Deaths / Cases

